---
title: "Temporal dynamics of Gahagan bifaces"
author: "Robert Z. Selden, Jr."
date: "January 31, 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the oxcAAR package

This document includes supplemental materials for paper: "A quantitative assessment of intraspecific morphological variation in Gahagan bifaces." Further information regarding the radiocarbon dates discussed below can be found in the text of the article.

```{r install}
if(!require('devtools')) install.packages('devtools')
library(devtools)
install_github('ISAAKiel/oxcAAR')
# load oxcAAR and OxCal
library(oxcAAR)
quickSetupOxcal()
```

## Calibration

### Calibration and plots for dates from Burial Pit 2 at the Gahagan Mound site

Three AMS dates (UGA12296, ISGS A0465, and ISGS A0466) were reported from Burial Pit 2 at the Gahagan Mound site. More information about these dates can be found in the text, and in those works cited in the text of the article.

```{r gahagancal, echo=TRUE}
gahagan<-data.frame(bp=c(1000,950,910),
                    std=c(40,40,35),
                    names=c("UGA-12296","ISGS-A0466","ISGS-A0465")
)
gahagancal<-oxcalCalibrate(gahagan$bp,gahagan$std,gahagan$names)
gahagancal
plot(gahagancal)
calcurve_plot(gahagancal)
```

### Calibration and plots for the dates from F119 and F134 at the George C. Davis site

Two dates (Tx-913 and Tx-1206) were reported from Features 119 and 134, respectively. More information about these dates can be found in the text, and in those works cited in the text of the article.

```{r GCDcal, echo=TRUE}
gcd<-data.frame(bp=c(1150,986),
                std=c(80,90),
                names=c("Tx-913","Tx-1206")
)
gcdcal<-oxcalCalibrate(gcd$bp,gcd$std,gcd$names)
gcdcal
plot(gcdcal)
calcurve_plot(gcdcal)
```

### Calibration and plots for dates from Burial Pit 5 at the Mounds Plantation site

Three dates (Tx-55, Tx-56, and M-1466) were reported from Burial Pit 5 at the Mounds Plantation site. More information about these dates can be found in the text, and in those works cited in the text of the article.

```{r mplcal, echo=TRUE}
moundspl<-data.frame(bp=c(860,900),
                     std=c(120,100),
                     names=c("Tx-55","M-1466")
)
moundsplcal<-oxcalCalibrate(moundspl$bp,moundspl$std,moundspl$names)
moundsplcal
plot(moundsplcal)
calcurve_plot(moundsplcal)
```

### Calibration and plots for dates from AU2 at the J. B. White site

Seventeen dates (Beta-176624, UGA-12500, UGA-12509, UGA-12497, UGA-12498, UGA-12505, UGA-12495, Beta-176625, Beta-176623, UGA-12494, UGA-12503, UGA-12499, Beta-176620, Beta-176627, Beta-176628, UGA-12493, Beta-176630) were reported from AU2 at the J. B. White site. More information about these dates can be found in the text, and in those works cited in the text of the article.

```{r jbwcal, echo=TRUE}
jbw<-data.frame(bp=c(1190,1090,1090,1050,1030,990,980,980,970,970,950,920,910,890,880,860,710),
                     std=c(50,40,40,40,40,40,40,40,40,40,40,40,40,40,40,40,40),
                     names=c("Beta-176624","UGA-12500","UGA-12509","UGA-12497","UGA-12498", "UGA-12505","UGA-12495","Beta-176625","Beta-176623","UGA-12494","UGA-12503","UGA-12499","Beta-176620","Beta-176627","Beta-176628","UGA-12493","Beta-176630")
)
jbwcal<-oxcalCalibrate(jbw$bp,jbw$std,jbw$names)
jbwcal
plot(jbwcal)
calcurve_plot(jbwcal)
```

### Calibration and plots for the dates from the Baker site

Three dates (D-AMS 035170, D-AMS 035169, D-AMS 035290) were reported from Features 119 and 134, respectively. More information about these dates can be found in the text, and in those works cited in the text of the article.

```{r bkrcal, echo=TRUE}
bkr<-data.frame(bp=c(667,648,624),
                std=c(24,22,28),
                names=c("D-AMS 035170","D-AMS 035169","D-AMS 035290")
)
bkrcal<-oxcalCalibrate(bkr$bp,bkr$std,bkr$names)
bkrcal
plot(bkrcal)
calcurve_plot(bkrcal)
```

### Chronological model for Caddo contexts yielding Gahagan bifaces

For reasons explained in the text, Tx-1206 and Tx-56 are included in the model as outliers. The script that appears below achieves the same results in R and in OxCal. While Figure 3 in the article was produced using this script in OxCal 4.3.2 for aesthetic purposes, the chronological model was created using the oxcAAR package.

```{r chronmod, echo=TRUE}
knitr::opts_chunk$set(cache=TRUE)
bgahagan <- 'Plot()
 {
  Phase("Contexts yielding Gahagan Bifaces")
  {
   Sequence("Gahagan Bifaces")
   {
    Boundary("start Gahagan Bifaces SCA");
    Phase("southern Caddo area")
    {
     R_Date("Tx-913", 1150, 80);
     R_Date("UGA-12296", 1000, 40);
     R_Date("Tx-1206", 986, 90)
     {
      Outlier();
     };
     R_Date("ISGS-A0466", 950, 40);
     R_Date("ISGS-A0465", 910, 35);
     R_Combine("Log 1")
     {
      R_Date("M-1466", 900, 100);
      R_Date("Tx-55", 860, 120);
     };
     R_Date("Tx-56", 475, 100)
     {
      Outlier();
     };
    };
    Boundary("end Gahagan bifaces SCA");
    Boundary("start Gahagan bifaces CTX");
    Phase("central Texas")
    {
     R_Date("Beta-176624", 1190, 50);
     R_Date("UGA-12500", 1090, 40);
     R_Date("UGA-12509", 1090, 40);
     R_Date("UGA-12497", 1050, 40);
     R_Date("UGA-12498", 1030, 40);
     R_Date("UGA-12505", 990, 40);
     R_Date("UGA-12495", 980, 40);
     R_Date("Beta-176625", 980, 40);
     R_Date("Beta-176623", 970, 40);
     R_Date("UGA-12494", 970, 40);
     R_Date("UGA-12503", 950, 40);
     R_Date("UGA-12499", 920, 40);
     R_Date("Beta-176620", 910, 40);
     R_Date("Beta-176627", 890, 40);
     R_Date("Beta-176628", 880, 40);
     R_Date("UGA-12493", 860, 40);
     R_Date("Beta-176630", 710, 40);
     R_Date("D-AMS 035170", 667, 24);
     R_Date("D-AMS 035169", 648, 22);
     R_Date("D-AMS 035290", 624, 28);
    };
    Boundary("end Ghagan Bifaces CTX");
   };
  };
 };
'
  
  bgahagan_results <- executeOxcalScript(bgahagan) 
  bgahagan_text <- readOxcalOutput(bgahagan_results)
  bgahagan_result_data <- parseOxcalOutput(bgahagan_text, only.R_Date = F)
  str(bgahagan_result_data)
  print(bgahagan_result_data)
  plot(bgahagan_result_data)
  # end of script
```

### Acknowledgments

Thanks to Martin Hinz for fielding questions related to the oxcAAR package, and to Derek Hamilton for his guidance with the chronological model.
